---
title: "Model Checking"
author: Owen G. Ward
date: 04/17/2023
format: 
  # html:
  #   toc: true
  #   toc-location: left
  pdf
bibliography: refs.bib
---


```{r setup}
#| include: false

knitr::opts_chunk$set(echo = FALSE, message = FALSE)

library(tidyverse)
library(RcppRoll)
library(cmdstanr)
library(posterior)
library(bayesplot)
library(loo)
library(here)

options(mc.cores = parallel::detectCores())

theme_set(theme_bw())


prob_positive <- function(stan_draws, param = "beta") {
  all_draws <- as_draws_df(stan_draws)
  ndraws <- nrow(all_draws)
  as_tibble(all_draws) %>% 
  select(starts_with(param)) %>% 
  apply(2, function(x) sum(x>0)/ndraws)
}

```


## Model Comparison

We want to compare these potential models, which is very similar
to the model we fit previously.
The only change is that we now need the history for
both players, not just the focal player.

```{r load in the initial data}
#| cache: true


lichess_data <- readRDS("../../rdata/lichess_pilot.RData")

bullet_60 <- lichess_data %>% 
  filter(Event == "Rated Bullet game") %>% 
  filter(TimeControl == "60+0") %>% 
  mutate(game_id = row_number())

get_hist <- function(user, games, prev_n) {
  hist_games <- games %>% 
    filter(White == user | Black == user) %>% 
    arrange(UTCDate, UTCTime) %>% 
    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
    select(White:BlackElo, focal_white, game_id) %>% 
    mutate(focal_result = case_when(
      (focal_white == 1 & Result == "1-0") ~ 1,
      (focal_white == 0 & Result == "0-1") ~ 1,
      (Result == "1/2-1/2") ~ 0.5,
      .default = 0
      )) %>% 
    mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
                              roll_mean(focal_result, n = prev_n)))
  hist_games
}


focal_players <- head(sort(table(c(bullet_60$Username)), decreasing = TRUE), 18)
## if don't use all focal players here then removing data because they won't be
## involved

top_players <- names(focal_players)


## then get the opponents of these players

full_opponents <- bullet_60 %>% 
  filter(Username %in% top_players) %>% 
  select(White, Black) 

all_opps <- c(full_opponents$White, full_opponents$Black)

opponents <- unique(all_opps)


## then need to get the history of these players, but maybe need
## to store game id
## can just match using date and time I guess?
## need to modify the function then to account for this

get_hist_opp <- function(user, games, prev_n) {
  hist_games <- games %>% 
    filter(White == user | Black == user) %>% 
    arrange(UTCDate, UTCTime) %>% 
    mutate(opp_white = ifelse(user == White, 1, 0)) %>% 
    select(White:BlackElo, opp_white, game_id) %>% 
    mutate(opp_result = case_when(
      (opp_white == 1 & Result == "1-0") ~ 1,
      (opp_white == 0 & Result == "0-1") ~ 1,
      (Result == "1/2-1/2") ~ 0.5,
      .default = 0
      ))
  
  ngames <- nrow(hist_games)
  
  if(ngames < prev_n) {
    hist_games <- hist_games %>% 
    mutate(opp_win_prop = opp_result/(1:ngames))
  }
  else{
    hist_games <- hist_games %>% 
      mutate(opp_win_prop = c(cumsum(opp_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
                              roll_mean(opp_result, n = prev_n)))
  }
  hist_games$player <- user
  hist_games
}
## so now the history is the history corresponding to the name
## in the "player" column


remaining_opponents <- opponents[!(opponents %in% top_players)]

opp_hist <- map_dfr(remaining_opponents, get_hist_opp, bullet_60, prev_n = 5) %>% 
  as_tibble()

## so this gives the history of the opponents for each of their 
## games, can use this to then first get their previous result, etc,
## then join this with the other results

prev_opp <- opp_hist %>% 
  group_by(player) %>% 
  arrange(UTCDate, UTCTime) %>% 
  mutate(prev_opp_result = lag(opp_result, default = 0), 
         prev_opp_prop = lag(opp_win_prop, default = 0)) %>%
  select(game_id, player, prev_opp_result, prev_opp_prop)

tidy_games <- map_dfr(top_players, get_hist, bullet_60, prev_n = 5) %>% 
  as_tibble()


all <- tidy_games %>% 
  left_join(prev_opp, by = "game_id") %>% 
  mutate(focal = ifelse(focal_white == 1, White, Black))
## this looks more reasonable now
## this has the history of the opponent included, not the history 
## of the focal player yet


final_data <- all %>% 
  group_by(focal) %>% 
  arrange(UTCDate, UTCTime) %>% 
  mutate(prev_focal_result = lag(focal_result, default = 0),
         prev_focal_prop = lag(focal_win_prop, default = 0)) %>% 
  ungroup()


## then tidy this up for stan

```


This object `prev_game_both` should have the previous result of both players in
each game, if that game is in the dataset.
Similarly, `final_prev` has the games played by the focal players, along 
with their previous result and their opponents previous result, if that
game was in the original dataset used (all 60 second bullet games here).

Then convert this into data for stan.


```{r get stan data for simple model}

final_data

stan_prev <- final_data %>% 
  mutate(WhiteElo = as.numeric(WhiteElo), 
         BlackElo = as.numeric(BlackElo)) %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>%
  mutate(elo_diff = ifelse(focal_white == 1, 
                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
  mutate(focal_id = match(focal_user, top_players)) %>% 
  select(focal_user, focal_id, focal_white, 
         prev_focal_result, prev_opp_result, focal_result) 


stan_data <- list(N = nrow(stan_prev),
                  J = length(top_players),
                  y = stan_prev$focal_result,
                  id = stan_prev$focal_id,
                  focal_prev = stan_prev$prev_focal_result,
                  opp_prev = stan_prev$prev_opp_result)

```


```{r fit the simple stan model}
#| eval: false

## remember no effect for non focal players in this at all 
## at the moment

stan_file <- "../Comp1.stan"

mod <- cmdstan_model(stan_file)

fit1_simple <- mod$sample(data = stan_data,
                  seed = 123,
                  chains = 4,
                  parallel_chains = 4,
                  refresh = 100)

## this fits at least now...

fit1_simple$summary()
mcmc_hist(fit1_simple$draws(c("alpha", "beta")))

```



## Notes on these models

- How does the choice of focal player make a difference in these models?

```{r thinking about the choice of focal}

table(bullet_60$Username)


tidy_bullet <- bullet_60 %>% 
  select(White:BlackElo, Username, game_id) %>% 
  mutate(opponent = ifelse(Username == White, Black, White),
         focal_white = ifelse(Username == White, 1, 0)) 


all_players <- c(tidy_bullet$Username, tidy_bullet$opponent)
players <- unique(all_players)

## then just get all the history for this
all_hist <- map_dfr(players, get_hist_opp, bullet_60, prev_n = 5) %>% 
  as_tibble() %>% 
  select(game_id:player, UTCDate, UTCTime)

min_hist <- all_hist %>% select(game_id:player)

## this should have each game appear twice, for each of the two players
## involved, so need to figure out how to handle that best
## left join with tidy games matching player to focal and then
## to opponent, rename the columns as needed


tidy_stan <- tidy_bullet %>% 
  left_join(min_hist, by = c("game_id", "Username" = "player")) %>% 
  arrange(UTCDate, UTCTime) %>% 
  rename(focal_result = opp_result, focal_win_prop = opp_win_prop) %>% 
  left_join(min_hist, by = c("game_id", "opponent" = "player")) %>% 
  as_tibble() 

## this is all good, just need to lag both and add them in,
## and also get the ids for the stan code


## getting the overall current win probabilities for each player also, 
## up to and including the current game

player_avg <- all_hist %>% 
  group_by(player) %>% 
  arrange(UTCDate, UTCTime, .by_group = TRUE) %>%
  mutate(nrow = row_number()) %>% 
  mutate(curr_avg = cumsum(opp_result)/nrow) %>% 
  select(game_id, player, curr_avg)

global_avg <- all_hist %>% 
  group_by(player) %>% 
  summarise(total_avg = mean(opp_result))

## then just join this in also to use it


```



```{r getting the final data for potential stan model}


final_data <- tidy_stan %>% 
  left_join(player_avg, by = c("Username" = "player", "game_id")) %>% 
  rename(focal_avg = curr_avg) %>% 
  left_join(player_avg, by = c("opponent" = "player", "game_id")) %>% 
  rename(opp_avg = curr_avg) %>% 
  group_by(Username) %>% 
  arrange(UTCDate, UTCTime) %>% 
  mutate(focal_prev_result = lag(focal_result, default = 0),
         focal_runn_avg = lag(focal_avg, n = 2, default = 0)) %>% 
  ungroup() %>% 
  group_by(opponent) %>% 
  arrange(UTCDate, UTCTime) %>% 
  mutate(opp_prev_result = lag(opp_result, default = 0),
         opp_runn_avg = lag(opp_avg, n = 2, default = 0)) %>% 
  ungroup() %>% 
  mutate(focal_id = match(Username, players),
         opp_id = match(opponent, players)) %>% 
  filter(focal_result != 0.5) %>% 
  select(-focal_avg, - opp_avg, - opp_result, 
         - focal_win_prop, - opp_win_prop)

final_data
## remove some unused columns

norm_data <- final_data %>% 
  left_join(global_avg, by = c("Username" = "player")) %>% 
  rename(focal_global_avg = total_avg) %>% 
  left_join(global_avg, by = c("opponent" = "player")) %>% 
  rename(opp_global_avg = total_avg) %>% 
  mutate(focal_prev_result = focal_prev_result - focal_global_avg,
         focal_runn_avg = focal_runn_avg - focal_global_avg,
         opp_prev_result = opp_prev_result - opp_global_avg,
         opp_runn_avg = opp_runn_avg - opp_global_avg)

```

Now this matches nicely, contains their historic winning average up to 2 games
previously, along with their result in the previous game.

Note that this has not been standardized currently.


```{r}

norm_data


stan_data_all <- list(N = nrow(norm_data),
                      J = length(players),
                      y = norm_data$focal_result,
                      focal_id = norm_data$focal_id,
                      opp_id = norm_data$opp_id,
                      focal_prev = norm_data$focal_prev_result,
                      opp_prev = norm_data$opp_prev_result,
                      focal_avg = norm_data$focal_runn_avg,
                      opp_avg = norm_data$opp_runn_avg)


saveRDS(stan_data_all, file = "../cluster_scripts/stan_data.RDS")

stan_file <- "../Comp2.stan"

mod2 <- cmdstan_model(stan_file)

fit2 <- mod2$sample(data = stan_data_all,
                    seed = 123,
                    chains = 4,
                    parallel_chains = 4,
                    refresh = 100,
                    iter_sampling = 500,
                    iter_warmup = 100)


```


## Fitting this to all data

Fit this model to all data on the cluster, just to see the results.
The traceplots and R-hat for this model indicates it doesn't work well,
does not converge.

```{r}
stan_fit <- readRDS("../cluster_scripts/Cluster_stan.RDS")

mcmc_hist(stan_fit$draws(c("alpha[1]", "alpha[2]")))

r_hat <- rhat(stan_fit)

summary(r_hat)

which(r_hat > 1.15)
## just problems with mu and tau for convergence here, otherwise good?

```




## Comparing models on smaller data

Is there a small subset of players who play each other often that we can
use in the dataset that can be used to compare these models?


```{r}

bullet_60 %>% 
  filter(White %in% focal_players) %>% 
  filter(Black %in% focal_players)

## no games between any of the players used to scrape the data
## would maybe need to get more players then, to use this for a comparison

```


```{r}
library(chessR)


chessdotcom_leaders <- chessdotcom_leaderboard(game_type = "live_bullet")

top_bullet <- chessdotcom_leaders %>% top_n(10, wt = -rank) %>% 
  select(player_id, username)


# chessdotcom_hikaru_recent <- get_raw_chessdotcom(usernames = "Hikaru",
#                                                  year_month = c(202201:202205))


games_22 <- top_bullet$username %>% 
  map_dfr(~get_raw_chessdotcom(.x, year_month = c(202201:202203)))



top_games <- games_22 %>% 
  filter(White %in% top_bullet$username) %>% 
  filter(Black %in% top_bullet$username)


## get 500 games here then for 1 month, probably enough to fit an initial model
## most of these games bullet too, so could use that

top_games %>% 
  group_by(White, Black, TimeClass) %>% 
  tally() %>% 
  arrange(-n)


bullet_games <- top_games %>% 
  filter(TimeClass == "bullet")


bullet_top10 <- saveRDS(bullet_games, file = "chess_com_bullet_10.RDS")

```




